---
title: "Covid-19"
author: "Hend Aljobaily"
date: "September 19, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Working Directory #
setwd("C:/Users/Hend Aljobaily/Desktop/Covid-19 Project")
```


```{r load-packages, include=FALSE}
# Load ibrary #
library(leaflet) # interactive map
library(leaflet.extras) # interactive map
library(dplyr) # for %>%
library(rgdal)
library(ggplot2) # plots
library(zoo) # moving averages
library(markdown) # R markdown
```

```{r}
# Load ibrary #
#library(shiny) # interactive plots
#library(leaflet) # interactive map
#library(leaflet.extras) # interactive map
#library(dplyr) # for %>%
#library(rgdal)
#library(ggplot2) # plots
#library(zoo) # moving averages
#library(markdown) # R markdown
```
# Metric 1 #
#### Create a map showing US COVID-19 case incidence per 100,000 in the last 7 days. Also, find and use state-level population estimates. 
##### Note: Data obtained on September 3, 2020 

Data obtained from CDC website (https://covid.cdc.gov/covid-data-tracker/#cases)
population estimates, cases per 100,000 for 7 days, longitudes, & latitudes were added to the dataset

```{r}
#import data
coviddata <- read.csv("covid-19.csv", header=TRUE)
```


### LEAFLET MAP ###

```{r message = FALSE}
#this is to add a popup with information about each state when clicking of the circle
coviddata <- coviddata%>%mutate(popup_info=paste(State,"<br/>","Population:",Population,"<br/>",
                                                 "Total Number of Cases:",Total.Cases,"<br/>",
                                                 "Total Cases Per 100,000:", Case.Rate.per.100000,"<br/>",
                                                 "Total Number of Cases In The Last 7 Days:",
                                                 Cases.in.Last.7.Days,"<br/>",
                                                 "Total Cases Per 100,000 In The Last 7 Days:",
                                                 Case.Rate.per.100000.in.the.last.7.days))

attach(coviddata)

colors<-c("YlOrRd") #the higher the number of cases in the last 7 days, the darker the color
bins <- c(0, 500, 1000, 2000, 5000, 10000, 20000, 40000) #adjust the defult bins
palette<-colorBin(colors, coviddata$Cases.in.Last.7.Days, bins=bins) #color will be based on this


#Create interactive map with leaflet
map<-leaflet()%>%addTiles()%>%fitBounds(min(Longitude), min(Latitude),max(Longitude), max(Latitude))%>%
  addCircleMarkers(data=Cases.in.Last.7.Days, lat=Latitude, lng=Longitude, 
                   radius = sqrt(Case.Rate.per.100000.in.the.last.7.days),
                   popup = popup_info, color = palette(Cases.in.Last.7.Days))%>%
  addLegend("topright",      #add a legend for the tottal numder of cases in the last 7 days
            pal = palette, 
            values = Cases.in.Last.7.Days,
            title = "Total Number of Cases",
            opacity = 1)%>%
  addControl("COVID-19 Cases in The Last 7 Days", position = "bottomright")%>%   #add title to the map
  addControl("Note: The size of circle represents cases per 100,000 people in the last 7 days", 
             position = "bottomleft",  className="map-title")%>% # add note about the size of circles
  addControl("Note: Data obtained from CDC on Septermber 3, 2020", 
             position = "bottomleft",  className="map-title") # add date of when data obtained

map #plot map
```

<span style="color:red">**Click on the circle markers on the map for more information about state-level COVID-19 cases.**</span>

# Metric 2 #
#### Use state-level historical data to visualize:  
##### Note: Data obtained on September 4, 2020

```{r}
# import data
## data obtained from the COVID Tracking Project (https://covidtracking.com/)
histdata <- read.csv("all-states-history.csv", header=TRUE)
histdata$newdate <- as.character(strptime(as.character(histdata$date), "%Y%m%d")) #reformat date variable
histdata$newdate <- as.Date(histdata$newdate)

histdata <- histdata %>%  # 7-day moving average calculation
  dplyr::arrange(desc(State)) %>% 
  dplyr::group_by(State) %>% 
  dplyr::mutate(MA7 = zoo::rollmean(Cases.Per.100000, k = 7, fill = NA)) %>% # moving average for positive cases per 100,000
  dplyr::mutate(MA77 = zoo::rollmean(PCR.Per.100000, k = 7, fill = NA)) %>% # moving average for PCR conducted per 100,000
  dplyr::mutate(MA777 = zoo::rollmean(Positivity, k = 7, fill = NA)) %>% # moving average for positivity percentage
  dplyr::ungroup()

```

### SHINY PLOTS ####

#### Part A: ####
#### Visualize the 7-day moving average of new cases reported per 100,000 ####

```{r}
ui1 <- fluidPage(
  titlePanel(h3("7-Day Moving Average of New COVID-19 Cases Per 100,000 People",align = "center")),
  sidebarLayout(
    sidebarPanel(
      selectizeInput("stateInput", "U.S. State/Territory", choices=NULL, selected=NULL)
    ),
    mainPanel(plotOutput(outputId = "plot1", width  = "500px",height = "350px"))
  )
)

server1 <- function(input, output, session) {
  updateSelectizeInput(session, 'stateInput',
                       choices = sort(unique(histdata$State)),
                       server = TRUE
  )
  
  output$plot1 <- renderPlot({
    ggplot(histdata, aes(x = newdate, y = MA7))+
      geom_line(data=histdata[histdata$State==input$stateInput,],color = "#09557f") + 
      scale_x_date(date_breaks = "1 week", date_minor_breaks = "1 day", date_labels = "%d %B %Y",
                   limits = c(min(histdata$newdate), max = max(histdata$newdate)),
                   expand=c(0,0))+
      #limit decided using max(histdata$Cases.Per.100000)
      ylim(0,3600)+ 
      xlab("Date")+ylab("New COVID-19 Cases Per 100,000 People")+
      theme(axis.text.x = element_text(angle=40, hjust=1))
  })
}
shinyApp(ui1, server1)

```

#### Part B: ####
#### Visualize the 7-day moving average of daily PCR tests conducted per 100,000 people ####

```{r}
ui2 <- fluidPage(
  titlePanel(h3("7-Day Moving Average of Daily PCR Conducted Per 100,000 People",align = "center")),
  sidebarLayout(
    sidebarPanel(
      selectizeInput("stateInput", "U.S. State/Territory", choices=NULL, selected=NULL)
    ),
    mainPanel(plotOutput(outputId = "plot2", width  = "500px",height = "350px"))
  )
)

server2 <- function(input, output, session) {
  updateSelectizeInput(session, 'stateInput',
                       choices = sort(unique(histdata$State)),
                       server = TRUE
  )
  
  output$plot2 <- renderPlot({
    ggplot(histdata, aes(x = newdate, y = MA77))+
      geom_line(data=histdata[histdata$State==input$stateInput,],color = "#09557f") + 
      scale_x_date(date_breaks = "1 week", date_minor_breaks = "1 day", date_labels = "%d %B %Y",
                   limits = c(min(histdata$newdate), max = max(histdata$newdate)),
                   expand=c(0,0))+
      #limit decided using max(histdata$PCR.Per.100000)
      ylim(0,55000)+ 
      xlab("Date")+ylab("Daily PCR Conducted Per 100,000 People")+
      theme(axis.text.x = element_text(angle=40, hjust=1))
  })
}
shinyApp(ui2, server2)
```

#### Part C: ####
#### Visualize the 7-day moving average of percent positivity ####
##### (person-level positive PCR tests divided by the total number of people who received a PCR test each day) #####

```{r}

ui3 <- fluidPage(
  titlePanel(h3("7-Day Moving Average of COVID-19 Positivity Percentage",align = "center")),
  sidebarLayout(
    sidebarPanel(
      selectizeInput("stateInput", "U.S. State/Territory", choices=NULL, selected=NULL)
    ),
    mainPanel(plotOutput(outputId = "plot3", width  = "500px",height = "350px"))
  )
)

server3 <- function(input, output, session) {
  updateSelectizeInput(session, 'stateInput',
                       choices = sort(unique(histdata$State)),
                       server = TRUE
  )
  
  output$plot3 <- renderPlot({
    ggplot(histdata, aes(x = newdate, y = MA777))+
      geom_line(data=histdata[histdata$State==input$stateInput,],color = "#09557f") + 
      scale_x_date(date_breaks = "1 week", date_minor_breaks = "1 day", date_labels = "%d %B %Y",
                   limits = c(min(histdata$newdate), max = max(histdata$newdate)),
                   expand=c(0,0))+
      #limit decided using max(histdata$Positivity)
      ylim(0,1)+ 
      xlab("Date")+ylab("COVID-19 Positivity Percentage")+
      theme(axis.text.x = element_text(angle=40, hjust=1))
  })
}
shinyApp(ui3, server3)
```

### REGULAR PLOTS ###
#### These plots inlcude COVID-19 related information for Colorado Nebraska, New Mexico, Wyoming, Arizona, Oklahoma, Kansas, Utah, and Texas ####

```{r}
# Keep only desired states
st <- histdata %>% 
  filter(State %in% c("Colorado","Texas","Nebraska","New Mexico", 
                      "Wyoming","Arizona","Oklahoma","Kansas","Utah"))
```

#### Part A: ####
#### Visualize the 7-day moving average of new cases reported per 100,000 ####

```{r warning = FALSE}
st %>%
  ggplot(aes(x = newdate, y = MA7, color=State)) +
  geom_point() + geom_line()+
  scale_x_date(date_breaks = "1 week", date_minor_breaks = "1 day", date_labels = "%d %B %Y",
               expand=c(0,0))+
  #limit decided using max(histdata$Cases.Per.100000)
  ylim(0,3600)+ 
  xlab("Date")+ylab("New COVID-19 Cases Per 100,000 People")+
  theme(axis.text.x = element_text(angle=40, hjust=1))+
  ggtitle("7-Day Moving Average of New COVID-19 Cases Per 100,000 People")
```

#### Part B: ####
#### Visualize the 7-day moving average of daily PCR tests conducted per 100,000 people ####

```{r warning = FALSE}
st %>%
  ggplot(aes(x = newdate, y = MA77, color=State)) +
  geom_point() + geom_line()+
  scale_x_date(date_breaks = "1 week", date_minor_breaks = "1 day", date_labels = "%d %B %Y",
               expand=c(0,0))+
  #limit decided using max(histdata$PCR.Per.100000)
  ylim(0,55000)+ 
  xlab("Date")+ylab("Daily PCR Conducted Per 100,000 People")+
  theme(axis.text.x = element_text(angle=40, hjust=1))+
  ggtitle("7-Day Moving Average of Daily PCR Conducted Per 100,000 People")
```

#### Part C: ####
#### Visualize the 7-day moving average of percent positivity ####
##### (person-level positive PCR tests divided by the total number of people who received a PCR test each day) #####

```{r warning = FALSE}
st %>%
  ggplot(aes(x = newdate, y = MA777, color=State)) +
  geom_point() + geom_line()+
  scale_x_date(date_breaks = "1 week", date_minor_breaks = "1 day", date_labels = "%d %B %Y",
               expand=c(0,0))+
  #limit decided using max(histdata$Positivity)
  ylim(0,1)+ 
  xlab("Date")+ylab("COVID-19 Positivity Percentage")+
  theme(axis.text.x = element_text(angle=40, hjust=1))+
  ggtitle("7-Day Moving Average of COVID-19 Positivity Percentage")
```

